A Novel Model-Free Learning Approach for Dynamic Target Tracking Problem and its Validation using Virtual Robot Experimentation Platform

### 2020-06-30 paperMIC2020/paperIEEE-MIC2020-V1.tex
fixed bib

Tracking a random moving target using a mobile robot, for instance, is a challenging task. This is mainly due to their inherent complex nonlinear dynamics dynamics. Most of the target tracking algorithms proposed in the literature either rely on A) complex mathematical models, B) simplified mathematical models, and C) nonlinear control techniques or driven by large amount of mostly offline data that leads to an overwhelming degree of computational complexity.

Over the past years mobile robots has been used in several applications in commercial and military sectors such as surveillance, search and rescue missions,coverage optimization, cooperative localization and dynamic target tracking tasks. In all of these mentioned applications a fleet of robot i.e agents interact with each other to achieve a certain goal. Leader-Follower or dynamic target tracking problem has recieved an extensive amount of study and research in the world of cooperative control theory due to it's wide promising applications such as search and rescure missions, wildlife monitoring and survillance to name a few . In a typical leader follower formation a number of robots referred to as followers apply local control actions to follow leader's robots in a specific predefined path such as \hl {cyclic, circular motion and time varying communication topologies.} \hl {limitations of other methods} many control methods were proposed however they have had the following limitations:
\\(1) a static leader/target is used.
\\(2) the leade/target's location or dynamic model is predetermined.  
\\(3) expensive hardware platforms are used to achieve the mobile target tracking. 
 \\The main contribution of this work is the development of a model-free learning approach to control the follower robot by which it overcome many of the limitations mentiond above as it does not rely on any prior information of the mathmetical model or the dynamic model. The steering angle and the linear speed of the follower robot are determined by collecting the position and orientation of both the leader and the follower. This set of information is gathered online over a finite period of time. The optimal control actions are then generated by utilizing Bellman's principal of optimality which acts as model free-reinfocment learning approach that allows the follower robot to follow the path of the leader while oviding collision by maintaining a safe distance. In this paper the proposed algorithm is further validated using a commercially available robot simulator, CoppeliaSim. This paper acts as first milestone in generalizing the algorithm to solve more sophisticated problem such as coverage and mapping. \hl{cite area coverage papers}
 \\The rest of the paper is orgnized as follows. Section II lays down the problem setting of the leader follower problem and mathemetical models of the robots and the state error. The model free actor-critic reinforcment learning and its key steps are described in section III.Section IV illustrates computer simulations for different secnarios that reflects the effectivness of the proposed method followed by conlusion and future work presented in section V.

\todo[inline]{Take some references from the following paragraphs}

  Need to be rewritten some parts of~\cite{Mi2018-c1}
  % The task of localization and mapping in the context of leader-follower problems is very challenging. Such a task is becoming a ubiquitous  facet of modern life due to its promising application in addressing not only the leader-follower problem but also applicable in industry industry with heavy equipment trucks, in agriculture with autonomous crop maintenance, in home life with the autonomous vacuuming robot, along with research and design as seen in~\cite{Roberts2007,Durmus2015,Elara2014,HevrdejsKnoll2017-Indoor}.  In most leader-follower problem addressed in the literature to date, it is assumed that leader/target's position/state is known \textit{a priori}, which may not be the case in many applications such as tracking a mobile point source~\cite{Xu2013}. Furthermore, the localization and mapping task is of a paramount importance in addressing the problem of mobile target tracking, which has a number of promising applications, such as robotic navigation, search and rescue mission, wildlife monitoring, autonomous surveillance, to name a few. A large body of research has been conducted in the literature to address the localization and mapping problem. Relatively few papers have addressed the localization and mapping problem in the context of leader-follower tasks, where either (i) a static leader/target is used~\cite{Fernando2015}, (ii) the leader/target's location is predetermined, or (iii) expensive hardware platforms are used to implement the mobile target tracking strategy (see the work presented in~\cite{Yang2014}, for example). 
%
%The work done in this paper addresses some of the aforementioned issues by carrying out a cost-effective and easy-to-implement localization and mapping algorithm in the context of a leader--follower problem, a wheeled mobile robot is employed as a follower and a target moving on a two-dimensional (2D) plane is used as a leader. We emphasize that the position of the  leader is unknown to the follower robot \emph{a priori.} Therefore, the position and orientation of the follower robot and the position of the leader are to be simultaneously estimated while building the map of the environment of the robot using a set of networked wireless radio sensors placed on the ground. It is assumed that a radio sensor is mounted on the leader. The follower robot receives range measurements [herein the received signal strength indicator (RSSI) from radio sensors] from all radio sensors in its operating range. These measurements are then used to estimate the position and orientation of the follower robot, the 2D position of the leader, and the 2D positions of radio sensors placed in the robot's environment. Once the follower robot estimates its states (position and orientation) and the position of the leader, it moves towards the leader using the motion controller running onboard the follower robot. The design and implementation of the current work are based on the preliminary work of the robot navigation and mapping strategy conducted by the authors in~\cite{HevrdejsKnoll2017-Indoor,KnHeMi2017-c1,HeKnMi2017-c1}. Therefore, the main objective is to advance the work in~\cite{HevrdejsKnoll2017-Indoor,KnHeMi2017-c1,HeKnMi2017-c1} for a more general case, where a mobile robot is to follow a leader  whose position is unknown \emph{a priori.}   
%n}

@@

% Note that, the policy iteration computational setup rearranges the temporal difference expression~\eqref{eq:valueFunctionEstimated} such that %
 %
 \begin{equation}
 \label{eq:const}
 \mathbf{z}_k^T\, \mathbf{P}\, \mathbf{z}_k - \mathbf{z}_{k+1}^T\, \mathbf{P}\, \mathbf{z}_{k+1} = \mathbf{z}_k^T\, \bar{\mathbf{P}}\, \mathbf{z}_k. 
 \end{equation}
 %
 This equation is utilized repeatedly in order to evaluate a certain policy during at least $\eta\ge \bar n, \bar n= (2+2)(2+2+1)/2$ evaluation steps (i.e., the lowest evaluation interval spans $k$ to $k+\bar n$ calculation samples) in order to update the critic weights vector  $\boldsymbol{\omega}=\mathrm{vec}(\mathbf{P}),$ which consists of connection weights between the neurons of the hidden layer and the output layer of the critic neural network shown in Fig.~\eqref{fig:nnCritic}. The operator $\mathrm{vec}(\mathbf{P})$ forms the columns of the $\mathbf{P}$ matrix into a column vector $\mathbf{\omega}$ of dimension $\bar{n}=15$ since the matrix $\mathbf{P}$ is a symmetric matrix. The left hand side of~\eqref{eq:const} is expressed using the following critic approximation form %
 %
 $$\hat{V}(\mathbf{z}_k)-\hat{V}(\mathbf{z}_{k+1})=\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1}),$$
 %
 where $\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})=\bm{\rho}(\mathbf{z}_k)-\bm{\rho}(\mathbf{z}_{k+1}) \in \mathbb{R}^{10 \times 1}, \, \bm{\rho}(\mathbf{z}_k)=\left(\mathbf{z}^q_k\otimes\mathbf{z}^h_k\right)$ $(q=1,\dots, 5, \,\, h=q,\dots,5),$ and $\boldsymbol{\omega}^T = [0.5 \, P^{11},P^{12},P^{13},P^{14},  \, 0.5 \, P^{22}, P^{23},P^{24},$ $\,0.5 \, P^{33}, P^{34},\,0.5 \ P^{44}, ]^T \in \mathbb{R}^{1\times 15}$ ($P^{ij}$ is the $ij^{th}$ entry of matrix $\mathbf{P}$). %
 %
 The critic weights $\boldsymbol{\omega}$ are updated using a gradient descent approach, where the tuning error $\varepsilon_k$ at each computational instance $k$ follows $\varepsilon_k =\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})-{v}_k,$ where $v_\kappa = \frac{1}{2}\mathbf{z}_{k}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k}$. As detailed earlier, it is required to perform at least $\eta \ge \bar n$ evaluation steps before updating the critic weights $\boldsymbol{\omega}$ (i.e., finding the new improved policy). Hence, it is required to minimize the sum of square errors such that %
 %
 \begin{multline*}
   \delta_c =\sum_{\kappa=0}^{\eta-1}\frac{1}{2}(\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k+\kappa,k+\kappa+1})-{v}_{k+\kappa})^2 = \frac{1}{2}\| \mathbf{v} - \bm{\Lambda}\bm{\omega}\|^2\\
   =\frac{1}{2}\left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right)^T \left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right), 
 \end{multline*}
 where $\bm{\Lambda} = [{\bf o}_0,{\bf o}_1,\ldots,{\bf o}_{\eta-1}]^T \in \mathbb{R}^{\eta\times 15}$ with ${\bf o}_\kappa = \tilde{\bm{\rho}}^T({\bf z}_{k+\kappa, k+\kappa+1}) \in \mathbb{R}^{1\times 15}$ and ${\bf v} =[v_0,v_1,\ldots,v_{\eta-1}]^T \in \mathbb{R}^{\eta}$ with $v_\kappa = \frac{1}{2}\mathbf{z}_{k+\kappa}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k+\kappa}$ for $\kappa = 0,1,\ldots, \eta-1$. %
 Therefore, the update law of the critic weights using the gradient decent approach for at least $\bar n$ samples is given by %
 \begin{multline}
   \bm{\omega}^{[r+1]} = \bm{\omega}^{[r]} - \ell_c\frac{\partial\delta_c}{\partial \bm{\omega}} = \bm{\omega}^{[r]} - \ell_c\left(-\bm{\Lambda}^T\mathbf{v} + \bm{\Lambda}^T \bm{\Lambda}\bm{\omega}^{[r]}\right)\\ 
 =\bm{\omega}^{[r]} - \ell_c \bm{\Lambda}^T\left(\bm{\Lambda}\bm{\omega}^{[r]}-\mathbf{v}\right), 
 \label{eq:criticWeights}
 \end{multline}
 %
 where $0<\ell_c<1$ is a critic learning rate and $r$ is the  update index of the critic weights.

@@

 \bibliographystyle{IEEEtran}
 \bibliography{bib/refsSuruzWeb,bib/refsBooksTRTheses,bib/refsReinforcementLearningADP}
\bibliography{bib/refsSuruzWeb,bib/refsMultiAgent,bib/refsRoboticsJournals,bib/refsRoboticsConferences,bib/refsGenericControl,bib/refsBooksTRTheses,bib/refsReinforcementLearningADP,bib/refsRL-Keshtkar}
\end{document}

>>>>

Tracking a random moving target using a mobile robot, for instance, is a challenging task. This is mainly due to their inherent complex nonlinear dynamics. Most of the target tracking algorithms proposed in the literature either rely on A) complex mathematical models, B) simplified mathematical models, and C) nonlinear control techniques or driven by large amount of mostly offline data that leads to an overwhelming degree of computational complexity.

Over the past years mobile robots has been used in several applications in commercial and military sectors such as surveillance, search and rescue missions,coverage optimization, cooperative localization and dynamic target tracking tasks \cite{kolling2006},\cite{Encarnacao2001},\cite{Ju2001}.  In all of these mentioned applications a fleet of robot i.e agents interact with each other to achieve a certain goal. Leader-Follower or dynamic target tracking problem has recieved an extensive amount of study and research in the world of cooperative control theory due to it's wide promising applications such as search and rescure missions, wildlife monitoring and survillance to name a few . In a typical leader follower formation a number of robots referred to as followers apply local control actions to follow leader's robots in a specific predefined path such as \hl {cyclic, circular motion and time varying communication topologies.} \hl {limitations of other methods} many control methods were proposed however they have had the following limitations:
\\(1) a static leader/target is used.
\\(2) the leade/target's location or dynamic model is predetermined.  
\\(3) expensive hardware platforms are used to achieve the mobile target tracking.

The main contribution of this work is the development of a model-free learning approach to control the follower robot by which it overcome many of the limitations mentiond above as it does not rely on any prior information of the mathmetical model or the dynamic model. The steering angle and the linear speed of the follower robot are determined by collecting the position and orientation of both the leader and the follower. This set of information is gathered online over a finite period of time. The optimal control actions are then generated by utilizing Bellman's principal of optimality which acts as model free-reinfocment learning approach that allows the follower robot to follow the path of the leader while avoiding collision by maintaining a safe distance. In this paper the proposed algorithm is further validated using a commercially available robot simulator, CoppeliaSim. This paper acts as first milestone in generalizing the algorithm to solve more sophisticated problem such as coverage and mapping. \hl{cite area coverage papers}  
The rest of the paper is orgnized as follows. Section II lays down the problem setting of the leader follower problem and mathemetical models of the robots and the state error. The model free actor-critic reinforcment learning and its key steps are described in section III.Section IV illustrates computer simulations for different secnarios that reflects the effectivness of the proposed method followed by conlusion and future work presented in section V.

@@

% Note that, the policy iteration computational setup rearranges the temporal difference expression~\eqref{eq:valueFunctionEstimated} such that %
 %
 \begin{equation}
 \label{eq:const}
 \mathbf{z}_k^T\, \mathbf{P}\, \mathbf{z}_k - \mathbf{z}_{k+1}^T\, \mathbf{P}\, \mathbf{z}_{k+1} = \mathbf{z}_k^T\, \bar{\mathbf{P}}\, \mathbf{z}_k. 
 \end{equation}
 %
 This equation is utilized repeatedly in order to evaluate a certain policy during at least $\eta\ge \bar n, \bar n= (2+2)(2+2+1)/2$ evaluation steps (i.e., the lowest evaluation interval spans $k$ to $k+\bar n$ calculation samples) in order to update the critic weights vector  $\boldsymbol{\omega}=\mathrm{vec}(\mathbf{P}),$ which consists of connection weights between the neurons of the hidden layer and the output layer of the critic neural network shown in Fig.~\eqref{fig:nnCritic}. The operator $\mathrm{vec}(\mathbf{P})$ forms the columns of the $\mathbf{P}$ matrix into a column vector $\mathbf{\omega}$ of dimension $\bar{n}=15$ since the matrix $\mathbf{P}$ is a symmetric matrix. The left hand side of~\eqref{eq:const} is expressed using the following critic approximation form %
 %
 $$\hat{V}(\mathbf{z}_k)-\hat{V}(\mathbf{z}_{k+1})=\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1}),$$
 %
 where $\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})=\bm{\rho}(\mathbf{z}_k)-\bm{\rho}(\mathbf{z}_{k+1}) \in \mathbb{R}^{10 \times 1}, \, \bm{\rho}(\mathbf{z}_k)=\left(\mathbf{z}^q_k\otimes\mathbf{z}^h_k\right)$ $(q=1,\dots, 5, \,\, h=q,\dots,5),$ and $\boldsymbol{\omega}^T = [0.5 \, P^{11},P^{12},P^{13},P^{14},  \, 0.5 \, P^{22}, P^{23},P^{24},$ $\,0.5 \, P^{33}, P^{34},\,0.5 \ P^{44}, ]^T \in \mathbb{R}^{1\times 15}$ ($P^{ij}$ is the $ij^{th}$ entry of matrix $\mathbf{P}$). %
 %
 The critic weights $\boldsymbol{\omega}$ are updated using a gradient descent approach, where the tuning error $\varepsilon_k$ at each computational instance $k$ follows $\varepsilon_k =\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k,k+1})-{v}_k,$ where $v_\kappa = \frac{1}{2}\mathbf{z}_{k}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k}$. As detailed earlier, it is required to perform at least $\eta \ge \bar n$ evaluation steps before updating the critic weights $\boldsymbol{\omega}$ (i.e., finding the new improved policy). Hence, it is required to minimize the sum of square errors such that %
 %
 \begin{multline*}
   \delta_c =\sum_{\kappa=0}^{\eta-1}\frac{1}{2}(\boldsymbol{\omega}^T\tilde{\bm{\rho}}(\mathbf{z}_{k+\kappa,k+\kappa+1})-{v}_{k+\kappa})^2 = \frac{1}{2}\| \mathbf{v} - \bm{\Lambda}\bm{\omega}\|^2\\
   =\frac{1}{2}\left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right)^T \left(\mathbf{v} - \bm{\Lambda}\bm{\omega}\right), 
 \end{multline*}
 where $\bm{\Lambda} = [{\bf o}_0,{\bf o}_1,\ldots,{\bf o}_{\eta-1}]^T \in \mathbb{R}^{\eta\times 10}$ with ${\bf o}_\kappa = \tilde{\bm{\rho}}^T({\bf z}_{k+\kappa, k+\kappa+1}) \in \mathbb{R}^{1\times 15}$ and ${\bf v} =[v_0,v_1,\ldots,v_{\eta-1}]^T \in \mathbb{R}^{\eta}$ with $v_\kappa = \frac{1}{2}\mathbf{z}_{k+\kappa}^T \, \bar{\mathbf{P}} \, \mathbf{z}_{k+\kappa}$ for $\kappa = 0,1,\ldots, \eta-1$. %
 Therefore, the update law of the critic weights using the gradient decent approach for at least $\bar n$ samples is given by %
 \begin{multline}
   \bm{\omega}^{[r+1]} = \bm{\omega}^{[r]} - \ell_c\frac{\partial\delta_c}{\partial \bm{\omega}} = \bm{\omega}^{[r]} - \ell_c\left(-\bm{\Lambda}^T\mathbf{v} + \bm{\Lambda}^T \bm{\Lambda}\bm{\omega}^{[r]}\right)\\ 
 =\bm{\omega}^{[r]} - \ell_c \bm{\Lambda}^T\left(\bm{\Lambda}\bm{\omega}^{[r]}-\mathbf{v}\right), 
 \label{eq:criticWeights}
 \end{multline}
 %
 where $0<\ell_c<1$ is a critic learning rate and $r$ is the  update index of the critic weights.

@@

	\bibliographystyle{IEEEtran}
\bibliography{bib/refsSuruzWeb,bib/refsMultiAgent,bib/refsRoboticsJournals,bib/refsRoboticsConferences,bib/refsGenericControl,bib/refsBooksTRTheses,bib/refsReinforcementLearningADP,bib/refsRL-Keshtkar}
\end{document}
